### 2023-01-27 article.tex
Add weissgerber2018we

Here we define paradata as the data that describe the process of generating data.
In genetic epidemiology, the data generated are mostly the results 
of an analysis (e.g. predicting a person having a disease),
as done by computer code.
In such context, one could argue incorrectly that
it is the -usually English- scientific paper that best describes which computation steps 
have taken place.
However, this has the unrealistic 
assumption that there is a perfect match between the paper and the code.
In this chapter it is argued that the source code should should be supplied,
as these are the true paradata: if the paper and code disagree, it is the
code that has generated the results.
The chapter concludes by some rules how to better code to serve as paradata,
and hence allowing computational research to be truly reproducible.

@@

% Alexa and Blake?
Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that involve computation.
Both manuscripts are accepted by an equally prestigous journal after peer view. 
One researcher, however, does not supply the
computer code that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses the effect of supplying the computer 
code (from now on: 'code') 
and claims that code is paradata and that this paradata 
is just as important as the paper itself.

@@

\begin{verbatim}
We compared the values of x and y using a one-tailed Student's T-test,
as we expect the average value of x to be less than the average value of y.
\end{verbatim}

We ignore the choices of words and style of this sentence: what is
important is the content: a one-tailed T test is performed.
Taking a look at the (in this case, the programming language R) code, 
we find the following line:

@@

The line of code is simplified (as the result of the T test is never stored),
but the code is correct:
\verb|t.test| is indeed the name of an R function to do a T-test
and it is reasonable to assume that this code, when glancing over the code
of an article, is correct.

@@

@@

>>>>

Here we define paradata as data about the data collection process.
In genetic epidemiology, the data generated are mostly the results 
of an analysis (e.g. predicting a person having a disease),
as done by computer code.
In such context, one could argue incorrectly that
it is the scientific paper that best describes which computation steps 
have taken place.
However, this has the unrealistic 
assumption that there is a perfect match between the paper and the code.
In this chapter it is argued that the source code should be supplied,
as these are the true paradata: if the paper and code disagree, it is the
code that has generated the results.
The chapter concludes by some rules how to better code to serve as paradata,
and hence allowing computational research to be truly reproducible.

@@

% Alex and Blake?
Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that involve computation.
Both manuscripts are accepted by an equally prestigous journal after peer view. 
One researcher, however, does not supply the
computer code that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses the effect of supplying the computer 
code (from now on: 'code') 
and claims that code is paradata and that this paradata 
is just as important as the paper itself.

@@

\begin{verbatim}
We compared the values of x and y using an unpaired one-tailed Student's T-test,
as we expect the average value of x to be less than the average value of y.
\end{verbatim}

We ignore the choices of words and style of this sentence: what is
important is the content: an unpaired one-tailed T test is performed.
Taking a look at the (in this case, the programming language R) code, 
we find the following line:

@@

The line of code is simplified (as the result of the T test is never stored),
but the code is correct:
\verb|t.test| is indeed the name of an R function to do an unpaired T-test
and it is reasonable to assume that this code, when glancing over the code
of an article, is correct.

@@

To approach the same problem from a different angle:
correctly reporting the T-test in a reproducable ways
is already a problem, with only 80 out of a selected 179 papers
fully reporting the parity of T-test (i.e. paired or unpaired) 
\cite{weissgerber2018we}.
In those unresolved 99 papers, 
the code (if published) would clarify this.

@@

% Duran et al., grounds for trust
% [A] process is epistemically opaque relative to a cognitive agent X at time t
% just in case X does not know at t all of the epistemically relevant elements of
% the process (Humphreys 2009, 618)
% 
% On Computational reliability
% 
% (CR) if S’s believing p at t results from m, then S’s belief in p at t is justified.
% where S is a cognitive agent, p is any truth-valued proposition related to the
% results of a computer simulation, t is any given time, and m is a reliable com‑
% puter simulation.
%
%
% 4 Sources for Computational Reliabilism
%
% 1. Verification and validation methods
% 2. Robustness analysis for computer simulations
% 3. A history of (un)successful implementations
% 4. Expert knowledge
